---
title: "Introduction to fatalityanalysisr"
author: "Kena"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# An Introduction to fatalityanalysisr
fatalityanalysisr uses data from the US National Highway Traffic Safety Administration's Fatality Analysis Reporting System, which is a nationwide census providing the American public yearly data regarding fatal injuries suffered in motor vehicle traffic crashes. The package contains functions to identify the files, read them, analyze and summarize data and map the data points. 


```{r}
library(fatalityanalysisr)
sessionInfo()
```
##Make file name:
`make_filename()` creates a name for the accident csv file based on the year provided.

The source code of this function is:
```{r source_code_make_filename, eval=FALSE, tidy=TRUE}
make_filename <- function(year) {
  year <- as.integer(year)
  sprintf("accident_%d.csv.bz2", year)
}
```
Based on the year provided it creates the file name. 
 

```{r example, eval=TRUE, tidy=FALSE}
make_filename(2006)

```
##Read file name:
`fars_read()` reads in the US National Highway Traffic Safety Administration's Fatality Analysis Reporting System data.

The source code of this function is:
```{r source_code_fars_read, eval=FALSE, tidy=TRUE}
fars_read <- function(filename) {
  if (!file.exists(filename))
    stop("file '", filename, "' does not exist")
  data <- suppressMessages({
    readr::read_csv(filename, progress = FALSE)
  })
  dplyr::tbl_df(data)
}
```
Based on the name provided it reads the file. 
 
##Map state:
`fars_map_state()` will plot the accidents on a map for a given state and year.

The source code of this function is:
```{r source_code_map_state, eval=FALSE, tidy=TRUE}
fars_map_state <- function(state.num, year) {
  filename <- make_filename(year)
  data <- fars_read(filename)
  state.num <- as.integer(state.num)

  if(!(state.num %in% unique(data$STATE)))
    stop("invalid STATE number: ", state.num)
  data.sub <- dplyr::filter(data, STATE == state.num)
  if(nrow(data.sub) == 0L) {
    message("no accidents to plot")
    return(invisible(NULL))
  }
  is.na(data.sub$LONGITUD) <- data.sub$LONGITUD > 900
  is.na(data.sub$LATITUDE) <- data.sub$LATITUDE > 90
  with(data.sub, {
    maps::map("state", ylim = range(LATITUDE, na.rm = TRUE),
              xlim = range(LONGITUD, na.rm = TRUE))
    graphics::points(LONGITUD, LATITUDE, pch = 46)
  })
}

```
##Read years:
`fars_read()` will read in multiple Fatality Analysis Reporting System data files based on the years provided..

The source code of this function is:
```{r source_code_read_years, eval=FALSE, tidy=TRUE}
fars_read_years <- function(years) {
  lapply(years, function(year) {
    file <- make_filename(year)
    tryCatch({
      dat <- fars_read(file)
      dplyr::mutate(dat, year = year) %>%
        dplyr::select(MONTH, year)
    }, error = function(e) {
      warning("invalid year: ", year)
      return(NULL)
    })
  })
}

```
Based on the years provided, it reads the data. 


 

